The Revisited Hidden Weight Bit Function

### 2024-12-06 RevisedHiddenWeight.tex
Cleanup of tex code (pdf unchanged)

\def\anon{0} %% set to 1 for anonymous submissions, hides acknowledgements and author names
\def\full{1} %% set to 0 for springer proceedings

\ifnum\full=1
\documentclass[11pt]{llncs}

\addtolength{\parskip}{1pt}
\else
\documentclass[10pt, runningheads]{llncs}
\usepackage{times}

\fi

@@

\begin{document}
    \setcounter{page}{1} 
	\title{The Revisited Hidden Weight Bit Function}

	\titlerunning{The Revisited Hidden Weight Bit Function}
\author{
	Pierrick M\'eaux\inst{1},
	Tim Seuré\inst{1}, 
	Deng Tang\inst{2}
%	Pierrick M\'eaux\orcidID{0000-0001-5733-4341}\inst{1},
%	Tim Seuré\inst{1}, 
%	Deng Tang\orcidID{0000-0002-8373-9200}\inst{2}
}
\authorrunning{P.\ M\'eaux, T.\ Seur\'e, D.\ Tang}

\institute{
	University of Luxembourg, Luxembourg\\ \email{pierrick.meaux@uni.lu, tim.seure@uni.lu} 
	%\and
%		University of Luxembourg, Luxembourg\\ \email{pierrick.meaux@uni.lu}
	\and
	Shanghai Jiao Tong University, Shanghai, China\\
	\email{dengtang@sjtu.edu.cn}
}

	%----------------------------------------------------------------
	\maketitle

\institute{
	University of Luxembourg, Luxembourg\\
	\email{pierrick.meaux@uni.lu}		
}

@@

%\pmnote{short intro:

%HWBF introduced in x, interesting for its cryptographic criteria and easy implementation. cryptographic properties studied in y such as balancedness, algebraic immunity and nonlinearity. 
%New potential uses in a stream cipher, or modifications. Use: transciphering and works aiming at finding Boolean functions easy to evaluate homomorphically, input oblivious algorithms. 
%Nonlinearity a bit low for direct use.

%Contributions: new function as addition of HWBF and quadratic bent function. We show when the function is balanced, formalize and study its nonlinearity. Bounds on the nonlinearity of $f$, generalization to a family of function, and an algorithm for estimating the nonlinearity for $n$ up to $80$. Comparisons with other functions and practical values of the parameters for $n$ up to $20$.

%}

\iffalse
%prompt

The Hidden Weight Bit Function (\hwbf{}) has received a lot of attention since its introduction by Bryant
in $1991$~\cite{IEEE:Bryant91}, it has been considered as the simplest example of function with binary decision
diagram of exponential size~\cite{IEEE:Bryant91,TIA:BoLSW99}. 
The facility of implementing this function in different models of computation since it consists in computing the Hamming weight of the input and applying a simple linear function, and it relatively good cryptographic properties exhibited in~\cite{DAM:WCST14} (balancedness, nonlinearity, degree and algebraic immunity) made it an interesting starting point candidate to design the filter of some stream cipher constructions.

Recently, new developments in stream ciphers lead to reconsider this function.
The new potential uses of stream ciphers with filter functions on a larger number number of variables such as Hybrid Homomorphic Encryption~\cite{CCS:NaeLauVai11} (\textsf{HHE}) gave a regain of interest in these functions. \textsf{HHE} aims at finding Boolean functions efficient to evaluate as an input oblivious algorithm. 
For example various new binary stream ciphers have been designed since $2016$ such as Kreyvium~\cite{JOC:CCFLNP18}, FLIP~\cite{EC:MJSC16}, Rasta~\cite{C:Rasta}, FiLIP~\cite{INDO:MCJS19} and Dasta~\cite{TOSC:BeiLea20}. 
Currently, the designs with Boolean functions with more than one hundred variables as a filter are the ones leading to the \textsf{HHE} with best latency~\cite{INDO:HofMeaRic20,CCS:CDPP22,CIC:MeaParPei24}.
%cite FHE for MPC?

The nonlinearity of the \hwbf{} being low for a direct use as filter function, several generalizations have been considered recently, aiming at strengthening this parameter and keeping or improving the others cryptographic relevant parameters such as balancedness, algebraic degree and and algebraic immunity. In these generalizations we count the functions introduced in~\cite{IEEE:Carlet22,eprint:CarPar23,DAM:MeaOza24}. The later article considers some functions obtained by computing the Hamming weight of the input and then simply applying a quadratic Boolean function.

In this article we consider another function that is weightwise quadratic, obtained by \textsf{XOR}-ing the \hwbf{} and a bent quadratic function. Therefore, we exhibit a family of Boolean functions both easy to compute and with good parameters, improving over the ones exhibited in~\cite{DAM:MeaOza24}. Various of the techniques we introduce to determine the nonlinearity are of independent interest, and allow to bound the maximum of the absolute Walsh spectrum for a wider family of weightwise quadratic functions. 
Our contributions are the following:

\begin{itemize}
	\item In Section~\ref{sec:defAndBal}, we define the revisited \hwbf{} $f$, and study its balancedness. This part is done by studying the (restricted) Walsh transform of some quadratic functions over sets of fixed Hamming weight, called slices. A recursive relation between these values allows to determine for which values of $n$ the function is balanced.
	\item In Section~\ref{sec:dnk}, we link the maximum Walsh coefficients of the revisited \hwbf{} to the coefficients of generating functions and use techniques from complex analysis to prove lower bounds on the nonlinearity of $f$. 
	Differing from usual results in Boolean function used in cryptography, this allows us to determine a good enough bound on the nonlinearity for a larger family of functions, and for all even number of variables $n$.
	\item In Section~\ref{sec:expAndComp}, we explain techniques to get a better bound for bounded values of $n$, giving a tighter bound for the nonlinearity of $f$ up to $n=80$. In this parts with experiments, we compare the bounds we obtain and the real value of the nonlinearity between $f$ and other weightwise quadratic functions such as the majority function, the \hwbf{}, and the two main examples from~\cite{DAM:MeaOza24}. These results show a neat advantage for the revisited \hwbf{} relatively the nonlinearity compared to function with similar computation costs.
	\item In Section~\ref{sec:otherParameters} we analyze the other cryptographic relevant parameters of $f$, such as degree, algebraic immunity and fast algebraic immunity. Similarly to the comparison for the nonlinearity the revisited \hwbf{} performs better or similarly to the function with similar computation costs.

\end{itemize}

\fi

@@

\bigskip
In this article, we propose a new generalization of the \hwbf{}, a weightwise quadratic function constructed by \textsf{XOR}-ing the \hwbf{} with a bent quadratic function. This leads to a family of Boolean functions that are computationally efficient and feature superior cryptographic properties compared to previous constructions. 
Additionally, the techniques we introduce to analyze nonlinearity have broader applicability and provide bounds on the maximum absolute Walsh spectrum for a wider class of weightwise quadratic functions. Our contributions are as follows:
\begin{itemize}
	\item\emph{Balancedness analysis} (Section~\ref{sec:defAndBal}): We define the revisited \hwbf{} in $n$ variables, hereafter denoted by $f$, and analyze its balancedness. This involves studying the (restricted) Walsh transform of specific quadratic functions over sets of fixed Hamming weights (called slices). By establishing recursive relations for these values, we determine for which values of $n$ the function is balanced.

    \item \emph{Nonlinearity bounds} (Section~\ref{sec:dnk}):
    We relate the maximum Walsh coefficients of the revisited \hwbf{} to the coefficients of generating functions and employ complex analysis techniques to prove lower bounds on the nonlinearity of $f$. 
    Unlike most studies on Boolean functions used in cryptography, this approach yields strong nonlinearity bounds for a broader class of functions and for all even $n$.

    \item \emph{Experimental results and comparisons} (Section~\ref{sec:expAndComp}): 
    For bounded values of $n$, we employ specific techniques to refine the nonlinearity bounds of $f$, providing tighter estimates up to $n = 80$. Through experiments, we compare these bounds with actual nonlinearity values for $f$ and other weightwise quadratic functions, such as the majority function, the \hwbf{}, and the two main examples in~\cite{DAM:MeaOza24}. 
    Our results highlight the revisited \hwbf{}'s superior nonlinearity relative to functions with similar computational costs.

    \item \emph{Analysis of other cryptographic parameters} (Section~\ref{sec:otherParameters}):  
    We evaluate other cryptographically relevant parameters of $f$, including degree, algebraic immunity, and fast algebraic immunity. Similar to the nonlinearity analysis, the revisited \hwbf{} outperforms or matches other functions with comparable computational costs.
\end{itemize}

@@

%Accordingly, the Boolean hypercube is partitioned into $n+1$ slices where the elements have the same Hamming weight.

@@

\begin{definition}[Algebraic Normal Form, degree]\label{def:anf}
	We call Algebraic Normal Form of a Boolean function $f\in\BN$ its unique representation as an element of the ring $\F_2[x_1,\dots,x_n]/(x_1^2+x_1,\dots,x_n^2+x_n)$, and we represent it as follows:
	\[f(x_1,\dots,x_n)= \sum_{I \subseteq [1,n]} a_I \left( \prod_{i \in I} x_i \right), \]%x_1,\dots,x_n=\sum_{I \subseteq [1,n]} a_I x^I,\]
	where $a_I\in \F_2$. The (algebraic) degree of $f$ is defined by $\degg(f)=0$ in case $f$ is the null function, and $\degg(f)=\max\{ |I|:I\subseteq[1,n],a_I=1\}$ otherwise.
\end{definition}

@@

%\begin{definition}[Algebraic Immunity] \label{def:ai}
%	The algebraic immunity of a Boolean function $f\in \Bfn{n}$, denoted as $\AI(f)$, is defined as:
%	\[ \AI(f) = \min_{g \neq 0}\{ \degg(g) \; | \; fg = 0 \; \text{or} \; (f + 1)g = 0 \}{,} \]
%	where $\degg(g)$ is the algebraic degree of $g$.
%	The function $g$ is called an annihilator of $f$ (or $f + 1$). 
%	Additionally we denote $\AN(f) = \min_{g \neq 0}\{ \degg(g) \; | \; fg = 0\}$.
%\end{definition}

@@

\begin{definition}[Nonlinearity, \eg{} page 79 in \cite{Carlet20}]%and weightwise nonlinearity
	\label{def:nl}
	The nonlinearity $\NL(f)$ of a Boolean function $f\in \BN$ is the minimum Hamming distance between $f$ and all the affine functions in $\BN$:
	\[ \NL(f) = \min\{ \hd(f,g):g\in\BN,\degg(g)\leq 1\}{,} \]
	where the Hamming distance between $f$ and $g$ is defined as $\hd(f,g)=|\{x\in\F_2^n:f(x)\neq g(x)\}|$. Alternatively, the nonlinearity of $f\in\BN$ can also be defined in terms of its Walsh transform:
	\[ \NL(f) = 2^{n-1}- \frac{1}{2} \max_{a\in \F_2^n}{|\wt{f}(a)|}. \]
\end{definition}

@@

\bigskip
The algebraic degree of multiples of a Boolean function is a crucial quantity for various attacks on stream ciphers that utilize a Boolean function as a filter. 
The well-known algebraic attack~\cite{EC:CouMei03} and fast algebraic attack~\cite{C:Courtois03} on filtered \textsf{LFSR}s have motivated the study of cryptographic properties such as algebraic immunity and fast algebraic immunity, investigated in numerous works, \eg{} \cite{FSE:Armknecht04,C:HawRos04,ISC:JiaZhaWan13})). Nowadays, these parameters are systematically determined for any function considered as a filter.

@@

\begin{definition}[Fast algebraic immunity, \eg{} \cite{EC:ACGKMR06}]\label{def:FAI} The fast algebraic immunity of a Boolean function $f\in \Bfn{n}$ is defined as:
\[
    \FAI(f)=\min \{ 2 \AI(f), \min\{\degg(g)+ \degg(fg):g\in\BN,1\leq\degg(g)<\AI(f)\}\}
\]

\end{definition}

@@

Recall that the Boolean symmetric functions in $n$ variables are those that are constant on the slice $\Ekn{k}{n}$ for every $k\in [0,n]$. 
This class of functions has been thoroughly studied in the context of cryptography, see \eg{} \cite{IEEE:Carlet04,IEEE:CanVid05,INDO:BraPre05,DM:SarMai07,IEEE:QFLW09,IEEE:CheLu11,Latin:Meaux19,CCDS:Meaux21,IEEE:CarMea21}.
In this article, the symmetric functions that we consider will mainly be the slice indicator function and the majority function.
%\begin{definition}[Elementary symmetric functions]
%Let $i\in [0,n]$, the elementary symmetric function of degree $i$ in $n$ variables, denoted $\sigma_{i,n}$, is the function which ANF contains all monomials of degree $i$ and no monomials of other degrees. 
%\end{definition}
\begin{definition}[Slice indicator functions]\label{def:slice} The indicator function of the slice of weight $k\in[0,n]$ is the function $\phikn kn\in\BN$ defined by $\phikn kn(x)=1$ if and only if $\w(x)=k$.
\end{definition}

@@

\begin{definition}[Hidden Weight Bit Function]\label{def:hwbf}
    We call Hidden Weight Bit Function (\hwbf{}) the Boolean function $h\in\BN$ defined by:
	\[h(x)= \left\{ 
	\begin{array}{l l}
	0 & \quad \text{ if } \w(x) =0, \\
	x_k & \quad \text{ if } \w(x)=k.
	\end{array} \right. \]

\end{definition}

@@

%We summarize the known results on the cryptographic criteria of weightwise affine functions we will use in this article in the following property.

@@

\subsection{Krawtchouk polynomials}
We use Krawtchouk polynomials and some of their properties to prove one of our main results. We give the necessary definition here and refer to \eg{}~\cite{book:MacSlo78} for more details.

@@

\section{Revisited \hwbf{} and balancedness}\label{sec:defAndBal}
%\subsection{Restricted Walsh transform of $f$ in a recursive fashion}

@@

\begin{definition}[Revisited \hwbf{}]\label{def:revHWBF}
For an even integer $n\geq 0$, we call revisited \hwbf{} the Boolean function $f\in\BN$ defined as:
\[f(x)=\sum_{i=1}^{n/2}(x_i+1) x_{i+\frac{n}{2}} + \sum_{k=1}^nx_k\phikn{k}{n}(x).\]
\end{definition}

Since $f$ is the sum of a quadratic function and a weightwise affine function, it is a weightwise quadratic function. Using the formalism of Definition~\ref{def:wwdegd}, $f$ is the weightwise quadratic function defined by $f_0(x)=0$ and $f_k(x)=x_k+\sum_{i=1}^{n/2}(x_i+1) x_{i+n/2}$ for $k\in[1,n]$. This form allows to derive the restricted Walsh transform of $f$, which can be a useful tool to study the balancedness and bound the nonlinearity of a function.
%Maybe add examples, such as OM24, M24
%We study the restricted Walsh transform  of $f$ to bound its nonlinearity.
\begin{proposition}\label{prop:restrWT}
Let $n\geq 0$ be an even integer, and let $f\in\BN$ be the revisited \hwbf{}. Consider any binary vector $a\in\F_2^n$, and define $c=a+e_k+\sum_{i=n/2+1}^n e_i$. Then the restricted Walsh transform of $f$ satisfies:
\[\wtk{f}{k}(a)=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+n/2}+ c \cdot x}.\]
\end{proposition}
\begin{proof}
We compute:
%\begin{align*}
%	\wtk{f}{k}(a)&=\sum_{x \in \Ekn{k}{n}} (-1)^{f(x)+a \cdot x}
%	=\sum_{x \in \Ekn{k}{n}} (-1)^{x_k + \sum_{i=1}^{n/2} (x_i+1) x_{i+\frac{n}{2}}+ a \cdot x}\\
%	&=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+\frac{n}{2}}+ c \cdot x},
%\end{align*}
%\begin{align*}
\begin{align*}
    \wtk{f}{k}(a)&=\sum_{x \in \Ekn{k}{n}} (-1)^{f(x)+a \cdot x}\\
	&=\sum_{x \in \Ekn{k}{n}} (-1)^{x_k + \sum_{i=1}^{n/2} (x_i+1) x_{i+n/2}+ a \cdot x}\\
	&=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+n/2}+\left(x_k+\sum_{i=n/2+1}^{n}x_i+a\cdot x\right)}\\
    &=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+n/2}+ c \cdot x}.\qedhere
\end{align*}
\end{proof}

%\begin{align*}
%\wtk{f}{k}(a)&=\sum_{x \in \Ekn{k}{n}} (-1)^{f(x)+ax}\\
%&=\sum_{x \in \Ekn{k}{n}} (-1)^{x_k + \sum_{i=1}^{n/2} (x_i+1) x_{i+\frac{n}{2}}+ ax}\\
%&=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+\frac{n}{2}}+ (c)x},
%\end{align*}
%where $c$ is the $n$-length binary vector such that $c=a+e_k+\sum_{i=n/2+1}^n e_i$ where the $e_i$ denotes the canonical vector with “1” in position $i$.

@@

\begin{proposition}\label{prop:recursiveDnka}
Let $n\geq 0$ be an even integer. Then for all $k\in[0,n]$ and $a\in\F_2^n$, the following hold true:
\begin{itemize}
    \item We have $\Dnka 00a=1$ and $\Dnka 0ka=0$ if $k\neq 0$.
    \item If $n\geq 2$, then for $b=(a_1,\dots,a_{n-2})\in\F_2^{n-2}$, we have:
    \[\Dnka{n}{k}{a}=\Dnka{n-2}{k}{b}+\left((-1)^{a_{n-1}} + (-1)^{a_{n}}\right)\Dnka{n-2}{k-1}{b} + (-1)^{1+a_{n-1}+a_n} \Dnka{n-2}{k-2}{b}.\]
\end{itemize}
\end{proposition}

@@

%The limit case of $n=0$ yields that $\Dnka 0ka=1$ if $k=0$ and $\Dnka 0ka=0$ otherwise.

%With these notations we obtain a recursive relation for $\Dnka{n}{k}{a}$, and therefore for $\wtk{f}{k}(a)$, denoting $b=(a_1,\cdots,a_{n-2})$ and $y=(x_1,\cdots,x_{n-2})$, we get for any even $n\geq 2$:

@@

%In this analysis we considered the different cases depending on the values of $a_{n-1}$ and $a_n$, but the same reasoning applies with any $a_i, a_{i+1}$ where $i$ is odd.

@@

\begin{table}
	\small
	\centering
	\begin{tabular}{|c|c|c|c|}
		\hline
		$a$ & $(0,0)$ & $(0,1)$ & $(1,1)$ \\ \hline
		$\Dnka{2}{0}{a}$ &$1$&$1$ & $1$\\
		$\Dnka{2}{1}{a}$ &$2$& $0$ & $-2$\\
		$\Dnka{2}{2}{a}$ &$-1$&$1$ & $-1$\\ \hline
	\end{tabular}

	\mycaption{Values of $\Dnka{2}{k}{a}$ for $n=2$.}\label{tab:Dnka}
\end{table}

@@

\begin{proof}
	For every $k\in[1,n]$, we have by Proposition~\ref{prop:restrWT} that:
	%\[
	%\wtk{f}{k}(a)=\sum_{x \in \Ekn{k}{n}} (-1)^{f(x)+ax}=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+\frac{n}{2}}+ c \cdot x}=\Dnka{n}{k}{d},\]
	%where $d$ is the permutation of $c$ obtained by sending the first $n/2$ elements to the odd positions, and the $n/2$ last ones to the even positions. 
	\begin{align*}
		\wt{f}(a)&=\sum_{x \in \Ekn{k}{n}} (-1)^{f(x)+a\cdot x}\\
		&=\sum_{\pi(x) \in \Ekn{k}{n}} (-1)^{f(\pi(x))+a \cdot \pi(x)}\\
		&=\sum_{\pi(x) \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_{\pi(i)} x_{\pi(i+n/2)}+ x_{\pi(k)}+a \cdot \pi(x)+ \sum_{i=1}^{n/2} x_{\pi(i+n/2)}}\\
		&=\sum_{\pi(x) \in \Ekn{k}{n}}(-1)^{d_n(x)+x_{\pi(k)}+\pi^{-1}(a)\cdot x+\sum_{i=1}^{n/2}x_{2i}}\\
		&=\sum_{x\in \Ekn{k}{n}}(-1)^{d_n(x)+(b+\pi(e_k))\cdot x}\\
		&=\Dnka{n}{k}{b+\pi(e_k)}.
	\end{align*}
	We conclude by applying $\wtk{f}{k}(a)=1+\sum_{k=1}^n\wtk{f}{k}(a)$.
\end{proof}

@@

\begin{theorem}\label{th:bal}
	Let $n\geq 2$ be an even integer.
	Then the Walsh transform of the revisited \hwbf{} $f\in\BN$ at $0_n$ satisfies:
	\[
		\wt{f}(0_n)=\left\{\begin{array}{l l}
			0&\text{ if }n\equiv 0\bmod 4\\
			-2\binom{(n-2)/2}{(n-2)/4}&\text{ if }n\equiv 2\bmod 4.\\
		\end{array}\right.
	\]
	Accordingly, $f$ is balanced if and only if $n \equiv 0\bmod 4$.
\end{theorem}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%\pmnote{ The point for the remark was to recall that we use the notation "$| \cdot |$" for absolute value over the complex, since before we used it only to denote "cardinal" or absolute value over integers.}

%\begin{remark}
%    Property~\ref{property:cauchy_estimate} is typically employed to bound the absolute value of complex numbers. Naturally, the bound remains valid when the complex number is real, as will be the case in our application.
%\end{remark}

@@

\mycomment{One can also prove the following:

\begin{proposition}
    For a fixed integer $k\geq 1$ and sufficiently large $n\geq b(k)$, it holds that:
    \[
        \max_{a\in\F_2^n}|\Dnka nka|=|\Dnka nk{0_n}|=\left(-\frac12\right)^k\sum_{i=\lceil k/2\rceil}^{n/2}(-4)^i\binom{n/2}{i}\binom{i}{k-i}.
    \]
\end{proposition}

>>>>

\def\anon{0} % set to 1 for anonymous submissions, hides acknowledgements and author names
\def\full{1} % set to 0 for springer proceedings

\ifnum\full=1
	\documentclass[11pt]{llncs}
	\addtolength{\parskip}{1pt}
\else
	\documentclass[10pt, runningheads]{llncs}
	\usepackage{times}
\fi

@@

\begin{document}

\setcounter{page}{1}

\title{The Revisited Hidden Weight Bit Function}
\titlerunning{The Revisited Hidden Weight Bit Function}

\ifnum\anon=0
	\author{%
		Pierrick M\'eaux\inst{1},
		Tim Seuré\inst{1}, 
		Deng Tang\inst{2}
		%Pierrick M\'eaux\orcidID{0000-0001-5733-4341}\inst{1},
		%Tim Seuré\inst{1}, 
		%Deng Tang\orcidID{0000-0002-8373-9200}\inst{2}
	}
	\authorrunning{P.\ M\'eaux, T.\ Seur\'e, D.\ Tang}
	\institute{%
		University of Luxembourg, Luxembourg\\
		\email{pierrick.meaux@uni.lu, tim.seure@uni.lu} 
		\and
		Shanghai Jiao Tong University, Shanghai, China\\
		\email{dengtang@sjtu.edu.cn}
	}
\else
	\author{}
	\institute{}
\fi

\maketitle

@@

@@

\bigskip

In this article, we propose a new generalization of the \hwbf{}, a weightwise quadratic function constructed by \textsf{XOR}-ing the \hwbf{} with a bent quadratic function. This leads to a family of Boolean functions that are computationally efficient and feature superior cryptographic properties compared to previous constructions. 
Additionally, the techniques we introduce to analyze nonlinearity have broader applicability and provide bounds on the maximum absolute Walsh spectrum for a wider class of weightwise quadratic functions. Our contributions are as follows:

\begin{itemize}
	\item\emph{Balancedness analysis} (Section~\ref{sec:defAndBal}): We define the revisited \hwbf{} in $n$ variables, hereafter denoted by $f$, and analyze its balancedness. This involves studying the (restricted) Walsh transform of specific quadratic functions over sets of fixed Hamming weights (called slices). By establishing recursive relations for these values, we determine for which values of $n$ the function is balanced.
    \item \emph{Nonlinearity bounds} (Section~\ref{sec:dnk}):
    We relate the maximum Walsh coefficients of the revisited \hwbf{} to the coefficients of generating functions and employ complex analysis techniques to prove lower bounds on the nonlinearity of $f$. 
    Unlike most studies on Boolean functions used in cryptography, this approach yields strong nonlinearity bounds for a broader class of functions and for all even $n$.
    \item \emph{Experimental results and comparisons} (Section~\ref{sec:expAndComp}): 
    For bounded values of $n$, we employ specific techniques to refine the nonlinearity bounds of $f$, providing tighter estimates up to $n = 80$. Through experiments, we compare these bounds with actual nonlinearity values for $f$ and other weightwise quadratic functions, such as the majority function, the \hwbf{}, and the two main examples in~\cite{DAM:MeaOza24}. 
    Our results highlight the revisited \hwbf{}'s superior nonlinearity relative to functions with similar computational costs.
    \item \emph{Analysis of other cryptographic parameters} (Section~\ref{sec:otherParameters}):  
    We evaluate other cryptographically relevant parameters of $f$, including degree, algebraic immunity, and fast algebraic immunity. Similar to the nonlinearity analysis, the revisited \hwbf{} outperforms or matches other functions with comparable computational costs.
\end{itemize}

@@

@@

\begin{definition}[Algebraic Normal Form, degree]\label{def:anf}
	We call Algebraic Normal Form of a Boolean function $f\in\BN$ its unique representation as an element of the ring $\F_2[x_1,\dots,x_n]/(x_1^2+x_1,\dots,x_n^2+x_n)$, and we represent it as follows:
	\[
		f(x_1,\dots,x_n)= \sum_{I \subseteq [1,n]} a_I \left( \prod_{i \in I} x_i \right),
	\]
	where $a_I\in \F_2$. The (algebraic) degree of $f$ is defined by $\degg(f)=0$ in case $f$ is the null function, and $\degg(f)=\max\{ |I|:I\subseteq[1,n],a_I=1\}$ otherwise.
\end{definition}

@@

@@

\begin{definition}[Nonlinearity, \eg{} page 79 in \cite{Carlet20}]%and weightwise nonlinearity
	\label{def:nl}
	The nonlinearity $\NL(f)$ of a Boolean function $f\in \BN$ is the minimum Hamming distance between $f$ and all the affine functions in $\BN$:
	\[
		\NL(f) = \min\{ \hd(f,g):g\in\BN,\degg(g)\leq 1\},
	\]
	where the Hamming distance between $f$ and $g$ is defined as $\hd(f,g)=|\{x\in\F_2^n:f(x)\neq g(x)\}|$. Alternatively, the nonlinearity of $f\in\BN$ can also be defined in terms of its Walsh transform:
	\[
		\NL(f) = 2^{n-1}- \frac{1}{2} \max_{a\in \F_2^n}{|\wt{f}(a)|}.
	\]
\end{definition}

@@

\bigskip

The algebraic degree of multiples of a Boolean function is a crucial quantity for various attacks on stream ciphers that utilize a Boolean function as a filter. 
The well-known algebraic attack~\cite{EC:CouMei03} and fast algebraic attack~\cite{C:Courtois03} on filtered \textsf{LFSR}s have motivated the study of cryptographic properties such as algebraic immunity and fast algebraic immunity, investigated in numerous works, \eg{} \cite{FSE:Armknecht04,C:HawRos04,ISC:JiaZhaWan13})). Nowadays, these parameters are systematically determined for any function considered as a filter.

@@

\begin{definition}[Fast algebraic immunity, \eg{} \cite{EC:ACGKMR06}]\label{def:FAI}
	The fast algebraic immunity of a Boolean function $f\in \Bfn{n}$ is defined as:
	\[
		\FAI(f)=\min \{ 2 \AI(f), \min\{\degg(g)+ \degg(fg):g\in\BN,1\leq\degg(g)<\AI(f)\}\}.
	\]
\end{definition}

@@

Recall that the Boolean symmetric functions in $n$ variables are those that are constant on the slice $\Ekn{k}{n}$ for every $k\in [0,n]$. 
This class of functions has been thoroughly studied in the context of cryptography, see \eg{} \cite{IEEE:Carlet04,IEEE:CanVid05,INDO:BraPre05,DM:SarMai07,IEEE:QFLW09,IEEE:CheLu11,Latin:Meaux19,CCDS:Meaux21,IEEE:CarMea21}.
In this article, the symmetric functions that we consider will mainly be the slice indicator function and the majority function.

\begin{definition}[Slice indicator functions]\label{def:slice}
	The indicator function of the slice of weight $k\in[0,n]$ is the function $\phikn kn\in\BN$ defined by $\phikn kn(x)=1$ if and only if $\w(x)=k$.
\end{definition}

@@

\begin{definition}[Hidden Weight Bit Function]\label{def:hwbf}
    We call Hidden Weight Bit Function (\hwbf{}) the Boolean function $h\in\BN$ defined by:
	\[h(x)= \left\{ 
	\begin{array}{l l}
		0 & \quad \text{ if } \w(x) =0, \\
		x_k & \quad \text{ if } \w(x)=k.
	\end{array} \right. \] 
\end{definition}

@@

@@

\subsection{Krawtchouk polynomials}

We use Krawtchouk polynomials and some of their properties to prove one of our main results. We give the necessary definition here and refer to \eg{}~\cite{book:MacSlo78} for more details.

@@

\section{Revisited \hwbf{} and balancedness}\label{sec:defAndBal}

@@

\begin{definition}[Revisited \hwbf{}]\label{def:revHWBF}
	For an even integer $n\geq 0$, we call revisited \hwbf{} the Boolean function $f\in\BN$ defined as:
	\[
		f(x)=\sum_{i=1}^{n/2}(x_i+1) x_{i+\frac{n}{2}} + \sum_{k=1}^nx_k\phikn{k}{n}(x).
	\]
\end{definition}

Since $f$ is the sum of a quadratic function and a weightwise affine function, it is a weightwise quadratic function. Using the formalism of Definition~\ref{def:wwdegd}, $f$ is the weightwise quadratic function defined by $f_0(x)=0$ and $f_k(x)=x_k+\sum_{i=1}^{n/2}(x_i+1) x_{i+n/2}$ for $k\in[1,n]$. This form allows to derive the restricted Walsh transform of $f$, which can be a useful tool to study the balancedness and bound the nonlinearity of a function.

\begin{proposition}\label{prop:restrWT}
	Let $n\geq 0$ be an even integer, and let $f\in\BN$ be the revisited \hwbf{}. Consider any binary vector $a\in\F_2^n$, and define $c=a+e_k+\sum_{i=n/2+1}^n e_i$. Then the restricted Walsh transform of $f$ satisfies:
	\[
		\wtk{f}{k}(a)=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+n/2}+ c \cdot x}.
	\]
\end{proposition}

\begin{proof}
	We compute:
	\begin{align*}
		\wtk{f}{k}(a)&=\sum_{x \in \Ekn{k}{n}} (-1)^{f(x)+a \cdot x}\\
		&=\sum_{x \in \Ekn{k}{n}} (-1)^{x_k + \sum_{i=1}^{n/2} (x_i+1) x_{i+n/2}+ a \cdot x}\\
		&=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+n/2}+\left(x_k+\sum_{i=n/2+1}^{n}x_i+a\cdot x\right)}\\
		&=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+n/2}+ c \cdot x}.\qedhere
	\end{align*}
\end{proof}

@@

\begin{proposition}\label{prop:recursiveDnka}
Let $n\geq 0$ be an even integer. Then for all $k\in[0,n]$ and $a\in\F_2^n$, the following hold true:
	\begin{itemize}
		\item We have $\Dnka 00a=1$ and $\Dnka 0ka=0$ if $k\neq 0$.
		\item If $n\geq 2$, then for $b=(a_1,\dots,a_{n-2})\in\F_2^{n-2}$, we have:
		\[
			\Dnka{n}{k}{a}=\Dnka{n-2}{k}{b}+\left((-1)^{a_{n-1}} + (-1)^{a_{n}}\right)\Dnka{n-2}{k-1}{b} + (-1)^{1+a_{n-1}+a_n} \Dnka{n-2}{k-2}{b}.
		\]
	\end{itemize}
\end{proposition}

@@

@@

@@

\begin{table}
	\small
	\centering
	\begin{tabular}{|c|c|c|c|}
		\hline
		$a$ & $(0,0)$ & $(0,1)$ & $(1,1)$ \\ \hline
		$\Dnka{2}{0}{a}$ &$1$&$1$ & $1$\\
		$\Dnka{2}{1}{a}$ &$2$& $0$ & $-2$\\
		$\Dnka{2}{2}{a}$ &$-1$&$1$ & $-1$\\ \hline
	\end{tabular}
	\mycaption{Values of $\Dnka{2}{k}{a}$ for $n=2$.}\label{tab:Dnka}
\end{table}

@@

\begin{proof}
	For every $k\in[1,n]$, we have by Proposition~\ref{prop:restrWT} that:
	\begin{align*}
		\wt{f}(a)&=\sum_{x \in \Ekn{k}{n}} (-1)^{f(x)+a\cdot x}\\
		&=\sum_{\pi(x) \in \Ekn{k}{n}} (-1)^{f(\pi(x))+a \cdot \pi(x)}\\
		&=\sum_{\pi(x) \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_{\pi(i)} x_{\pi(i+n/2)}+ x_{\pi(k)}+a \cdot \pi(x)+ \sum_{i=1}^{n/2} x_{\pi(i+n/2)}}\\
		&=\sum_{\pi(x) \in \Ekn{k}{n}}(-1)^{d_n(x)+x_{\pi(k)}+\pi^{-1}(a)\cdot x+\sum_{i=1}^{n/2}x_{2i}}\\
		&=\sum_{x\in \Ekn{k}{n}}(-1)^{d_n(x)+(b+\pi(e_k))\cdot x}\\
		&=\Dnka{n}{k}{b+\pi(e_k)}.
	\end{align*}
	We conclude by applying $\wtk{f}{k}(a)=1+\sum_{k=1}^n\wtk{f}{k}(a)$.
\end{proof}

@@

\begin{theorem}\label{th:bal}
	Let $n\geq 2$ be an even integer.
	Then the Walsh transform of the revisited \hwbf{} $f\in\BN$ at $0_n$ satisfies:
	\[
		\wt{f}(0_n)=\left\{\begin{array}{ll}
			0&\text{ if }n\equiv 0\bmod 4\\
			-2\binom{(n-2)/2}{(n-2)/4}&\text{ if }n\equiv 2\bmod 4.\\
		\end{array}\right.
	\]
	Accordingly, $f$ is balanced if and only if $n \equiv 0\bmod 4$.
\end{theorem}

@@

@@

@@

\subsection{Generalization to a family of weightwise quadratic functions}\label{sec:general}

In this part, we generalize the results of Section~\ref{sec:Cauchy} to all weightwise quadratic functions $f$ such that for every $k\in [1,n]$, the function $f_k$ in Definition~\ref{def:wwdegd} has exactly $t$ quadratic terms with no shared variables.
